---
title: "Padilla_Pcom_RNAseq"
author: "MVI"
date: "2026-01-21"
output: html_document
---

```{r}
setwd("~/Desktop/Padilla_Gamino/PoritesCompressa_PadillaGamino")
```

```{r}
#to install arrayQualityMetrics, using Rswitch to go into R 4.2 bc it seems to not work on 4.5.2
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("arrayQualityMetrics")
#BiocManager::install("DESeq2")
#BiocManager::install("WGCNA")
#BiocManager::install("flashClust")
#install.packages("RColorBrewer")
#install.packages("gplots")
#install.packages("pheatmap")
#install.packages("vegan")
#install.packages("ggplot2")
#install.packages("ggrepel")
#install.packages("VennDiagram")
#install.packages("ggtext")
#install.packages("reshape2")
#install.packages("tidyselect")
#install.packages("plotly")
#install.packages("Hmisc")

```


```{r}
library("DESeq2")                   
#library("arrayQualityMetrics")      
library("dplyr")                  
library("RColorBrewer")             
library("gplots")                   
library("tidyverse")               
library("pheatmap")                
library("vegan")                    
library("ggplot2")                 
library("ggrepel") 
library("WGCNA")
library("flashClust")
library("VennDiagram")
#library("KOGMWU")
library("ggtext")
#library("ggpubr")
library("reshape2")
library("tidyselect")
library("devtools")
#library("cmprsk")
#library("car")
library("Hmisc")
library("plotly")
```

read in annotations file
```{r}
anno <- read.delim("./Porites_compressa_emapper.emapper.annotations.txt", sep="\t", header = TRUE)

#gene = anno %>%
#  mutate(gene_ID = ifelse(Preferred_name == '-', query, Preferred_name))
#View(gene)
#head(gene)
#gene2query <- gene[,c(1,22)]
#View(gene2query)

gene = anno %>%
mutate(gene_ID = ifelse(Preferred_name == '-' | is.na(Preferred_name), query, Preferred_name))%>%
  mutate(gene_ID = make.names(gene_ID, unique = TRUE))
gene2query <- gene[,c(1,22)]
```



```{r}
countData <- read.table("./Padilla_Pcom_counts.txt")
length(countData[,1])
#40465
head(countData)
newColNames = c("NB_F1_K5", "B_F1_K5", "NB_F2_44", "B_F2_44", "B_F3_44", "NB_F3_K5", "B_M1_K5", "B_M2_K5", "B_F4_K5", "B_F5_K5", "NB_F4_44", "NB_F5_44", "NB_F6_K5", "B_M3_K5", "B_F6_44")
colnames(countData)=paste(newColNames)
length(countData[,1])
#40465
head(countData)

totalCounts=colSums(countData)
barplot(totalCounts, ylab="Raw Counts", xlab="P.compressa samples")
#save as rawcounts_012126 as 6.5x4 landscape

min(totalCounts)
#2837556
max(totalCounts)
#8944911
```

```{r}
totalQs <- as.data.frame(row.names(countData))
colnames(totalQs)=paste("query")

gene = anno %>%
  right_join(totalQs) %>%
mutate(gene_ID = ifelse(Preferred_name == '-' | is.na(Preferred_name), query, Preferred_name))%>%
  mutate(gene_ID = make.names(gene_ID, unique = TRUE))
gene2query <- gene[,c(1,22)]
```


Open, inspect, and get count data for symbiont counts
```{r}
countData_s <- read.table("./Padilla_PcomSym_counts.txt")
head(countData_s)
length(countData_s[,1])
#20047
newColNames_s = c("NB_F1_K5", "B_F1_K5", "NB_F2_44", "B_F2_44", "B_F3_44", "NB_F3_K5", "B_M1_K5", "B_M2_K5", "B_F4_K5", "B_F5_K5", "NB_F4_44", "NB_F5_44", "NB_F6_K5", "B_M3_K5", "B_F6_44")
colnames(countData_s)=paste(newColNames_s)
length(countData_s[,1])
#20047
head(countData_s)

totalCounts_s=colSums(countData_s)
barplot(totalCounts_s, ylab="Raw Counts", xlab="P. compressa symbiont")
#save as rawcounts_symbionts_012126 as 11x4 landscape in Figure directory

min(totalCounts_s)
#32438
max(totalCounts_s)
#362232
```

Now return to the whole data set of the host analysis
Make treatment groups for a metadata file
```{r}
corals=c("NB_F1_K5", "B_F1_K5", "NB_F2_44", "B_F2_44", "B_F3_44", "NB_F3_K5", "B_M1_K5", "B_M2_K5", "B_F4_K5", "B_F5_K5", "NB_F4_44", "NB_F5_44", "NB_F6_K5", "B_M3_K5", "B_F6_44")
c=data.frame(corals)
coraldata <-c
coraldata

treat=c("NB_F_K5", "B_F_K5", "NB_F_44", "B_F_44", "B_F_44", "NB_F_K5", "B_M_K5", "B_M_K5", "B_F_K5", "B_F_K5", "NB_F_44", "NB_F_44", "NB_F_K5", "B_M_K5", "B_F_44")
g=data.frame(treat)
colData<- g
colData

reef=c("K5", "K5", "44", "44", "44", "K5", "K5", "K5", "K5", "K5", "44", "44", "K5", "K5", "44")
r=data.frame(reef)
reefData<- r
reefData

bleach = c("NB", "B", "NB", "B", "B", "NB", "B", "B", "B", "B", "NB", "NB", "NB", "B", "B")
h=data.frame(bleach)
bleachData<- h
bleachData

sex=c("F", "F", "F", "F", "F", "F", "M", "M", "F", "F", "F", "F", "F", "M", "F")
s=data.frame(sex)
sexData<- s
sexData

allData = cbind(coraldata, colData, reefData, bleachData, sexData)
allData
```

Run preliminary DESeq and outlier analysis
```{r}
dds<-DESeqDataSetFromMatrix(countData=countData, colData=colData, design=~treat)
dds<-DESeq(dds)

vsd.ge=assay(vst(dds))
rl=vst(dds)
e=ExpressionSet(assay(rl), AnnotatedDataFrame(as.data.frame(colData(rl))))
v = setwd("~/Desktop/Padilla_Gamino")
arrayQualityMetrics(e,outdir=v,intgroup=c("treat"),force=T)
```
No outliers!

so switch back to R 4.5 for remaining script

rerun deseq
```{r}
dds<-DESeqDataSetFromMatrix(countData=countData, colData=colData, design=~treat)
dds<-DESeq(dds)
```


- Now we are going to retrieve the rlog data, which will give us better transformation when size factors vary across samples
- regularized log transformation, transforms COUNT DATA to log2 scale to minimize differences between samples for rows with small counts
```{r}
rlog=rlogTransformation(dds, blind=TRUE) 
rld=assay(rlog)
head(rld)
hist(rld)
#not really normally distributed, kinda skewed to having lots of low count data
#saved as histogram_pcom_012126 5x5
```
- The assay function allows you to access the matrix-like data, so that you can view it bc head(rld) gives you just a summary
- The histogram above displays the frequency of each binned rld count value

Raw Sample Comparison

- Making a sample distance heatmap
- as.matrix attempts to turn its argument into a matrix
- dist computes and returns the distance matrix computed using the specified distance measure to compute distances between the rows of a data matrix rld
- Create a sample heatmap of the relatedness based on the count distance/difference between samples
```{r}
sampleDists <- as.matrix(dist(t(rld)))

heatmap.2(as.matrix(sampleDists), key=F, trace="none",
          col=colorpanel(100, "black", "white"),
          margin=c(10, 10))

#saved as sampledisthm_pcom as landscape with 6x6 dimensions
```

jump to PCA
```{r}
head(rld)
length(rld[,1])
#40465
```

```{r}
rld_t=t(rld)
#colnames(rld_t)
pca <- prcomp(rld_t,center = TRUE)
li <- pca$sdev^2 / sum(pca$sdev^2)
pc1v <- round(li[1] * 100, 1)
pc2v <- round(li[2] * 100, 1)
pca_s <- as.data.frame(pca$x)
pca_s <- pca_s[,c(1,2)]

pca_s$Samples = row.names(pca_s)
pca_s$treat=allData$treat

pca_s$reef=allData$reef
pca_s$bleach=allData$bleach
pca_s$sex=allData$sex

head(pca_s)
```

```{r}
reef_col <- c("red3", "blue")
sex_shape = c(25, 24)
bleach_fill <- c("white", "white", "white", "red3", "blue")
bleach_fill_new <- c("white", "red3")

pca_plot <- ggplot(pca_s, aes(x=PC1, y=PC2, color = reef, pch = sex, fill = treat, group=treat)) +
  geom_point(size=3, alpha=0.75) +
  #  geom_text_repel(aes(label=Samples)) +
  scale_colour_manual(values=reef_col)+
  scale_fill_manual(values=bleach_fill, guide=guide_legend(override.aes = list(color=list("white","white","white","black","black"))))+
 scale_shape_manual(values=sex_shape) +
  #stat_ellipse()+
  #labs(color="reef", pch="sex", fill="bleach")+
  xlab(paste0("PC1: ",pc1v,"% variance")) +
  ylab(paste0("PC2: ",pc2v,"% variance")) +
  theme(panel.background = element_rect(color="black", fill="white"),
        axis.text = element_text(colour="black", size=10), 
        axis.title.x = ggtext::element_markdown(face="bold", size=12),
        axis.title.y = ggtext::element_markdown(face="bold",size=12),
        panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
        legend.position="right",
        legend.background=element_rect(color="black"),
        legend.text=element_text(face="bold", color="black", size=8),
        legend.title= element_blank(),
        legend.text.align = 0,
        legend.key.size=unit(10,"point"))
ggplotly(pca_plot)
#padilla_PCA_all 6x4
```


```{r}
#####separate by sex to do only female
rld_fem <- rld[,c(1:6,9:13,15)]

rld_fem_t=t(rld_fem)
#colnames(rld_t)
pca_fem <- prcomp(rld_fem_t,center = TRUE)
li_fem <- pca_fem$sdev^2 / sum(pca_fem$sdev^2)
pc1v_fem <- round(li_fem[1] * 100, 1)
pc2v_fem <- round(li_fem[2] * 100, 1)
pca_s_fem <- as.data.frame(pca_fem$x)
pca_s_fem <- pca_s_fem[,c(1,2)]

head(allData)
allData_fem <- subset(allData, sex=="F")

pca_s_fem$Samples = row.names(pca_s_fem)
pca_s_fem$treat=allData_fem$treat

pca_s_fem$reef=allData_fem$reef
pca_s_fem$bleach=allData_fem$bleach
#pca_s$sex=allData$sex

head(pca_s_fem)

reef_col <- c("red3", "blue")
#sex_shape = c(21, 24)
bleach_fill_fem <- c("white", "white", "red3", "blue")

pca_plot_fem <- ggplot(pca_s_fem, aes(x=PC1, y=PC2, color = reef, fill = treat, group=treat)) +
  geom_point(size=3, alpha=0.75, shape=21) +
  #  geom_text_repel(aes(label=Samples)) +
  scale_colour_manual(values=reef_col)+
  scale_fill_manual(values=bleach_fill_fem, guide=guide_legend(override.aes = list(color=list("white","white", "black","black"))))+
 #scale_shape_manual(values=reef_shape) +
  #stat_ellipse()+
  #labs(color="reef", pch="sex", fill="bleach")+
  xlab(paste0("PC1: ",pc1v_fem,"% variance")) +
  ylab(paste0("PC2: ",pc2v_fem,"% variance")) +
  theme(panel.background = element_rect(color="black", fill="white"),
        axis.text = element_text(colour="black", size=10), 
        axis.title.x = ggtext::element_markdown(face="bold", size=12),
        axis.title.y = ggtext::element_markdown(face="bold",size=12),
        panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
        legend.position="right",
        legend.background=element_rect(color="black"),
        legend.text=element_text(face="bold", color="black", size=8),
        legend.title= element_blank(),
        legend.text.align = 0,
        legend.key.size=unit(10,"point"))
ggplotly(pca_plot_fem)
#padilla_PCA_fem 6x4
```

```{r}
install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(pairwiseAdonis)

# Effect of reef
adonis2(pca$x ~ reef, data = pca_s, method='eu', na.rm = TRUE)
# Pr(>F) = 0.28

# Effect of bleach status
adonis2(pca$x ~ bleach, data = pca_s, method='eu', na.rm = TRUE)
# Pr(>F)=0.69

# Effect of sex
adonis2(pca$x ~ sex, data = pca_s, method='eu', na.rm = TRUE)
# Pr(>F)=0.18

# Effect of treat
adonis2(pca$x ~ treat, data = pca_s, method='eu', na.rm = TRUE)
# Pr(>F)=0.197

# Interaction between reef and sex
adonis2(pca$x ~ reef*sex, data = pca_s, method='eu', na.rm = TRUE)
#Pr(>F>=0.092 .)

# Interaction between reef and bleach
adonis2(pca$x ~ reef*bleach, data = pca_s, method='eu', na.rm = TRUE)
#Pr(>F>=0.49)

# Interaction between sex and bleach
adonis2(pca$x ~ sex*bleach, data = pca_s, method='eu', na.rm = TRUE)
#Pr(>F>=0.459)

# Interaction between reef, sex and bleach
adonis2(pca$x ~ reef*sex*bleach, data = pca_s, method='eu', na.rm = TRUE)
#Pr(>F>=0.182)

# Pairwise effect of treatment (to get individual pairwise comparisons)
pairwise.adonis2(pca$x ~ treat, data = pca_s, method='eu', na.rm = TRUE)
#none are significant

```

```{r}
# Effect of reef
adonis2(pca_fem$x ~ reef, data = pca_s_fem, method='eu', na.rm = TRUE)
# Pr(>F) = 0.077 .

# Effect of bleach status
adonis2(pca_fem$x ~ bleach, data = pca_s_fem, method='eu', na.rm = TRUE)
# Pr(>F)=0.684

# Effect of treat
adonis2(pca_fem$x ~ treat, data = pca_s_fem, method='eu', na.rm = TRUE)
# Pr(>F)=0.14

# Interaction between reef and bleach
adonis2(pca_fem$x ~ reef*bleach, data = pca_s_fem, method='eu', na.rm = TRUE)
#Pr(>F>=0.134)

# Pairwise effect of treatment (to get individual pairwise comparisons)
pairwise.adonis2(pca_fem$x ~ treat, data = pca_s_fem, method='eu', na.rm = TRUE)
#none are significant
```


i think i'm going to have to separate the data by sex for the plasticity. Can really only compare the females because there are so few males and they're all bleached and all from one reef
```{r}
source("https://raw.githubusercontent.com/seabove7/RandomFun/main/Plasticity_function/PCAplast_function.R")

rld_plast <- rld_fem
head(rld_plast)
rld_plast_t <- t(rld_plast)
pca_df_plast <- prcomp(rld_plast_t, center=TRUE)

allData_fem2 <- remove_rownames(allData_fem)
allData_fem2 <- column_to_rownames(allData_fem2, var="corals")
head(allData_fem2)

write.csv(allData_fem2, file="allData_fem2.csv", quote=F, row.names=TRUE)
allData_fem2 <- read.csv("allData_fem2.csv")
head(allData_fem2)
allData_fem2 <- as.data.frame(allData_fem2)

head(pca_df_plast)

plast_out_all <- PCAplast(pca = pca_df_plast, # the PCA dataframe containing the PCA eigenvalues
         data = allData_fem2, # the condition/treatment data corresponding to samples
         #sample_ID = "X", # the name of column that provide unique ID per sample (if blank, will pull rownames for this)
        num_pca =  2, # the number of PCAs to include in analysis (default is 'all', but you can specify another number with a minimum of 2 PCAs)
         control_lvl = "NB", # control level of the treatment. If blank, a control mean per control level is assumed
         group = "reef", # the grouping column (i.e., colony). If blank, will assume control level grouping only! was "symstat" in first submission; changed to Clone here
         control_col = "bleach") # what the 'treatment' column is called

plast_out_all
```

```{r}
plast_col = c("red3", "blue")
#"gray40","gray40","gray40","gray40","#CC6600","#CC6600","#CC6600","#CC6600","#CC6600","gray40", "gray40", "gray40", "gray40", "gray40", "#CC6600", "#CC6600", "#CC6600", "#CC6600", "#CC6600", "#CC6600", "gray40", "gray40", "gray40", "gray40", "gray40", "#CC6600", "#CC6600", "#CC6600", "#CC6600", "#CC6600"

#plast_reorder <- plast_out_all[c(1:4, 10:14, 21:25, 5:9, 15:20, 26:30),]
#plast_reorder$treat  = factor(plast_reorder$treat, levels=plast_reorder$treat)

plast_plot <- ggplot(data = plast_out_all, aes(x = reef, y = dist, group=treat)) + 
  geom_point(color=c("blue", "red3", "red3", "blue", "blue", "red3"), stroke=1, position = position_jitter(width = 0.2),aes(fill=reef)) +
 stat_summary(fun.data = "mean_sdl", fun.args = list(mult = 1), geom = "errorbar", width = 0, colour = c("black")) +
  stat_summary(fun = "mean", size = 0.8, shape=21, colour = "black", fill="gray") +
  scale_fill_manual(values=plast_col, guide=guide_legend(override.aes = list(color=list("red3", "blue"))))+
  #scale_x_discrete(limits = c("P0A","P2A", "RA", "P0S", "P2S", "RS"))+
  #scale_shape_manual(values=c(16, 17, 18)) +
  #ylim(0,30)+
  ylab("Plasticity")+
  xlab("Reef")+
  #labs(fill="Symbiosis State")+
  theme(panel.background = element_rect(color="black", fill="white"),
        axis.text = element_text(colour="black", size=10), 
        axis.title.x = ggtext::element_markdown(face="bold", size=12),
        axis.title.y = ggtext::element_markdown(face="bold",size=12),
        panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
        legend.position="right",
        legend.background=element_rect(color="black"),
        legend.text=element_text(face="bold", color="black", size=8),
        legend.title= element_blank(),
        legend.text.align = 0,
        legend.key.size=unit(10,"point"))
plast_plot
#saved as plasticity_fem_020226 as portrait with 4x3

summary(plast_out_all)
plast_aov <- aov(dist~reef, data=plast_out_all)
summary(plast_aov)

#             Df Sum Sq Mean Sq F value Pr(>F)
#reef         1  941.3   941.3   1.718   0.26
#Residuals    4 2191.1   547.8     


```

Make comparison variables and visualize the results

```{r}
#Look at results in dds for female K5 B vs NB
head(dds)
resK5_BF_NBF<- results(dds, contrast = c("treat", "B_F_K5", "NB_F_K5"))
resK5_BF_NBF
head(resK5_BF_NBF)
table(resK5_BF_NBF$padj<0.1)
# FALSE=25002 TRUE=229
summary(resK5_BF_NBF, alpha=0.05)
# out of 40465 with nonzero total read count
#adjusted p-value < 0.05
#LFC > 0 (up)       : 34, 0.084%
#LFC < 0 (down)     : 101, 0.25%
#outliers [1]       : 351, 0.87%
#low counts [2]     : 14883, 37%

#Look at results in dds for female 44 B vs NB
res44_BF_NBF<- results(dds, contrast = c("treat", "B_F_44", "NB_F_44"))
res44_BF_NBF
head(res44_BF_NBF)
table(res44_BF_NBF$padj<0.1)
# FALSE=23648 TRUE=48
summary(res44_BF_NBF, alpha=0.05)
#out of 40465 with nonzero total read count
#adjusted p-value < 0.05
#LFC > 0 (up)       : 8, 0.02%
#LFC < 0 (down)     : 15, 0.037%
#outliers [1]       : 351, 0.87%
#low counts [2]     : 16418, 41%

##a lot of low counts

#Look at results in dds for K5 female bleached versus male bleached
resK5_BM_BF<- results(dds, contrast = c("treat", "B_M_K5", "B_F_K5"))
resK5_BM_BF
head(resK5_BM_BF)
table(resK5_BM_BF$padj<0.1)
# FALSE=23279 TRUE=417
summary(resK5_BM_BF, alpha=0.05)
#out of 40465 with nonzero total read count
#adjusted p-value < 0.05
#LFC > 0 (up)       : 123, 0.3%
#LFC < 0 (down)     : 59, 0.15%
#outliers [1]       : 351, 0.87%
#low counts [2]     : 16418, 41%
```

Dispersion plot
```{r}
plotDispEsts(dds, main="Dispersion P. com")
```

to do other, bigger grouped comparisons (like all bleached versus all non-bleached / all females versus all males / all K5 vs 44) i'll need to re-dds based on a different column (sex or bleach or reef)

sex
```{r}
dds_s<-DESeqDataSetFromMatrix(countData=countData, colData=sexData, design=~sex)
dds_s<-DESeq(dds_s)

rlog_s=rlogTransformation(dds_s, blind=TRUE) 
rld_s=assay(rlog_s)
head(rld_s)
hist(rld_s)

#Look at results in dds_s for all male to female
head(dds_s)
res_M_F<- results(dds_s, contrast = c("sex", "M", "F"))
res_M_F
head(res_M_F)
table(res_M_F$padj<0.1)
# 26627    69 
summary(res_M_F, alpha=0.05)
# out of 40449 with nonzero total read count
#adjusted p-value < 0.05
#LFC > 0 (up)       : 23, 0.057%
#LFC < 0 (down)     : 19, 0.047%
#outliers [1]       : 450, 1.1%
#low counts [2]     : 13319, 33%
```

reef
```{r}
dds_r<-DESeqDataSetFromMatrix(countData=countData, colData=reefData, design=~reef)
dds_r<-DESeq(dds_r)

rlog_r=rlogTransformation(dds_r, blind=TRUE) 
rld_r=assay(rlog_r)
head(rld_r)
hist(rld_r)

#Look at results in dds_s for all KF to 44
head(dds_r)
res_K5_44<- results(dds_r, contrast = c("reef", "K5", "44"))
res_K5_44
head(res_K5_44)
table(res_K5_44$padj<0.1)
# 26715    70 
summary(res_K5_44, alpha=0.05)
# out of 40449 with nonzero total read count
#adjusted p-value < 0.05
#LFC > 0 (up)       : 23, 0.057%
#LFC < 0 (down)     : 23, 0.057%
#outliers [1]       : 417, 1%
#low counts [2]     : 13263, 33%
```

bleach
```{r}
dds_b<-DESeqDataSetFromMatrix(countData=countData, colData=bleachData, design=~bleach)
dds_b<-DESeq(dds_b)

rlog_b=rlogTransformation(dds_b, blind=TRUE) 
rld_b=assay(rlog_b)
head(rld_b)
hist(rld_b)

#Look at results in dds_s for all bleach to nonbleach
head(dds_b)
res_B_NB<- results(dds_b, contrast = c("bleach", "B", "NB"))
res_B_NB
head(res_B_NB)
table(res_B_NB$padj<0.1)
# 28324    58 
summary(res_B_NB, alpha=0.05)
# out of 40449 with nonzero total read count
#adjusted p-value < 0.05
#LFC > 0 (up)       : 3, 0.0074%
#LFC < 0 (down)     : 28, 0.069%
#outliers [1]       : 343, 0.85%
#low counts [2]     : 11740, 29%
```

those comparisons aren't very informative, so i'll just leave them for now and come back if i want to, but for now, going to continue with the original dds

create GO input files for
resK5_BF_NBF
res44_BF_NBF
resK5_BM_BF
```{r}
#resK5_BF_NBF
write.table(resK5_BF_NBF, file="./K5_BF_NBF_DE_26.txt", quote=F, sep="\t")
head(read.delim("./K5_BF_NBF_DE_26.txt"))
res_K5BFNBF<- read.delim("./K5_BF_NBF_DE_26.txt")
res_K5BFNBF = res_K5BFNBF %>%
  tibble::rownames_to_column(var = "query")
head(res_K5BFNBF)
nrow(res_K5BFNBF)
#40465

go_input_K5BFNBF = res_K5BFNBF %>%
  mutate(mutated_p = -log(pvalue)) %>%
  mutate(mutated_p_updown = ifelse(log2FoldChange < 0, mutated_p*-1, mutated_p*1)) %>%
  na.omit() %>%
  select(query, mutated_p_updown)
nrow(go_input_K5BFNBF)
#25231
head(go_input_K5BFNBF)
colnames(go_input_K5BFNBF) <- c("gene", "pval")
head(go_input_K5BFNBF)
write.csv(go_input_K5BFNBF, file="./K5BFNBF_GO.csv", quote=F, row.names=FALSE)

#res44_BF_NBF
write.table(res44_BF_NBF, file="./44_BF_NBF_DE_26.txt", quote=F, sep="\t")
head(read.delim("./44_BF_NBF_DE_26.txt"))
res_44BFNBF<- read.delim("./44_BF_NBF_DE_26.txt")
res_44BFNBF = res_44BFNBF %>%
  tibble::rownames_to_column(var = "query")
head(res_44BFNBF)
nrow(res_44BFNBF)
#40465
res_44BFNBF_05 <- as.data.frame(res_44BFNBF[which(res_44BFNBF$padj < 0.05), ])

go_input_44BFNBF = res_44BFNBF %>%
  mutate(mutated_p = -log(pvalue)) %>%
  mutate(mutated_p_updown = ifelse(log2FoldChange < 0, mutated_p*-1, mutated_p*1)) %>%
  na.omit() %>%
  select(query, mutated_p_updown)
nrow(go_input_44BFNBF)
#23696
head(go_input_44BFNBF)
colnames(go_input_44BFNBF) <- c("gene", "pval")
head(go_input_44BFNBF)
write.csv(go_input_44BFNBF, file="./44BFNBF_GO.csv", quote=F, row.names=FALSE)

#resK5_BM_BF
write.table(resK5_BM_BF, file="./K5_BM_BF_DE_26.txt", quote=F, sep="\t")
head(read.delim("./K5_BM_BF_DE_26.txt"))
res_K5BMBF<- read.delim("./K5_BM_BF_DE_26.txt")
res_K5BMBF = res_K5BMBF %>%
  tibble::rownames_to_column(var = "query")
head(res_K5BMBF)
nrow(res_K5BMBF)
#40465
res_K5BMBF_05 <- as.data.frame(res_K5BMBF[which(res_K5BMBF$padj < 0.05), ])

go_input_K5BMBF = res_K5BMBF %>%
  mutate(mutated_p = -log(pvalue)) %>%
  mutate(mutated_p_updown = ifelse(log2FoldChange < 0, mutated_p*-1, mutated_p*1)) %>%
  na.omit() %>%
  select(query, mutated_p_updown)
nrow(go_input_K5BMBF)
#23696
head(go_input_K5BMBF)
colnames(go_input_K5BMBF) <- c("gene", "pval")
head(go_input_K5BMBF)
write.csv(go_input_K5BMBF, file="./K5BMBF_GO.csv", quote=F, row.names=FALSE)
```

run GO
```{r}
#K5_BF_NBF
input="K5BFNBF_GO.csv"
goAnnotations="Porites_compressa_iso2go.tab"
goDatabase="go.obo"
goDivision="BP"
source("gomwu.functions.R")

gomwuStats(input, goDatabase, goAnnotations, goDivision,
           perlPath="perl",
           largest=0.1,
           smallest=5,
           clusterCutHeight=0.25)
#261
resultsBP=gomwuPlot(input,goAnnotations,goDivision,
                    absValue=1,
                    level1=0.1,
                    level2=0.05,
                    level3=0.01,
                    txtsize=1.5,
                    treeHeight=0.5,
)
#saved as BP_apoP0C_051425 as portrait with 45x10 dimensions in Figures


```

Come back to the other less informative comparisons
```{r}
#B_NB
write.table(res_B_NB, file="./B_NB_DE_26.txt", quote=F, sep="\t")
head(read.delim("./B_NB_DE_26.txt"))
res_BNB<- read.delim("./B_NB_DE_26.txt")
res_BNB = res_BNB %>%
  tibble::rownames_to_column(var = "query")
head(res_BNB)
nrow(res_BNB)
go_input_BNB = res_BNB %>%
  mutate(mutated_p = -log(pvalue)) %>%
  mutate(mutated_p_updown = ifelse(log2FoldChange < 0, mutated_p*-1, mutated_p*1)) %>%
  na.omit() %>%
  select(query, mutated_p_updown)
nrow(go_input_BNB)
#28382
head(go_input_BNB)
colnames(go_input_BNB) <- c("gene", "pval")
head(go_input_BNB)
write.csv(go_input_BNB, file="./BNB_GO.csv", quote=F, row.names=FALSE)

#res_K5_44
write.table(res_K5_44, file="./K5_44_DE_26.txt", quote=F, sep="\t")
head(read.delim("./K5_44_DE_26.txt"))
res_K544<- read.delim("./K5_44_DE_26.txt")
res_K544 = res_K544 %>%
  tibble::rownames_to_column(var = "query")
head(res_K544)
nrow(res_K544)
go_input_K544 = res_K544 %>%
  mutate(mutated_p = -log(pvalue)) %>%
  mutate(mutated_p_updown = ifelse(log2FoldChange < 0, mutated_p*-1, mutated_p*1)) %>%
  na.omit() %>%
  select(query, mutated_p_updown)
nrow(go_input_K544)
#26785
head(go_input_K544)
colnames(go_input_K544) <- c("gene", "pval")
head(go_input_K544)
write.csv(go_input_K544, file="./K544_GO.csv", quote=F, row.names=FALSE)

#res_M_F
write.table(res_M_F, file="./M_F_DE_26.txt", quote=F, sep="\t")
head(read.delim("./M_F_DE_26.txt"))
res_MF<- read.delim("./M_F_DE_26.txt")
res_MF = res_MF %>%
tibble::rownames_to_column(var = "query")
head(res_MF)
nrow(res_MF)
go_input_MF = res_MF %>%
  mutate(mutated_p = -log(pvalue)) %>%
  mutate(mutated_p_updown = ifelse(log2FoldChange < 0, mutated_p*-1, mutated_p*1)) %>%
  na.omit() %>%
  select(query, mutated_p_updown)
nrow(go_input_MF)
#26696
head(go_input_MF)
colnames(go_input_MF) <- c("gene", "pval")
write.csv(go_input_MF, file="./MF_GO.csv", quote=F, row.names=FALSE)
#1776105370681 in thank god doc
```

GO on these broad comparisons
```{r}
input="BNB_GO.csv"
goAnnotations="Porites_compressa_iso2go.tab"
goDatabase="go.obo"
goDivision="BP"
source("gomwu.functions.R")
gomwuStats(input, goDatabase, goAnnotations, goDivision,
           perlPath="perl",
           largest=0.1,
           smallest=5,
           clusterCutHeight=0.25)
#168
resultsBP=gomwuPlot(input,goAnnotations,goDivision,
                    absValue=1,
                    level1=0.1,
                    level2=0.05,
                    level3=0.01,
                    txtsize=1.5,
                    treeHeight=0.5,
)

input="K544_GO.csv"
goAnnotations="Porites_compressa_iso2go.tab"
goDatabase="go.obo"
goDivision="BP"
source("gomwu.functions.R")
gomwuStats(input, goDatabase, goAnnotations, goDivision,
           perlPath="perl",
           largest=0.1,
           smallest=5,
           clusterCutHeight=0.25)
#724
resultsBP=gomwuPlot(input,goAnnotations,goDivision,
                    absValue=1,
                    level1=0.1,
                    level2=0.05,
                    level3=0.01,
                    txtsize=1.5,
                    treeHeight=0.5,
)

input="MF_GO.csv"
goAnnotations="Porites_compressa_iso2go.tab"
goDatabase="go.obo"
goDivision="BP"
source("gomwu.functions.R")
gomwuStats(input, goDatabase, goAnnotations, goDivision,
           perlPath="perl",
           largest=0.1,
           smallest=5,
           clusterCutHeight=0.25)
#1042
resultsBP=gomwuPlot(input,goAnnotations,goDivision,
                    absValue=1,
                    level1=0.1,
                    level2=0.05,
                    level3=0.01,
                    txtsize=1.5,
                    treeHeight=0.5,
)
```

Visualizing DEGs for each comparison in volcano plots
```{r}
#K5_BF_NBF
# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
resK5_BF_NBF_df <- as.data.frame(resK5_BF_NBF)
head(resK5_BF_NBF_df)
resK5_BF_NBF_df <- resK5_BF_NBF_df %>%
  rownames_to_column(var="query") %>%
  left_join(gene2query)
head(resK5_BF_NBF_df)
resK5_BF_NBF_df$diffexpressed <- "NO"
head(resK5_BF_NBF_df)
# if log2Foldchange > 0 and padj < 0.05, set as "UP"
resK5_BF_NBF_df$diffexpressed[resK5_BF_NBF_df$log2FoldChange > 0 & resK5_BF_NBF_df$padj < 0.05] <- "UP"
# if log2Foldchange < 0 and padj < 0.05, set as "DOWN"
resK5_BF_NBF_df$diffexpressed[resK5_BF_NBF_df$log2FoldChange < 0 & resK5_BF_NBF_df$padj < 0.05] <- "DOWN"
head(resK5_BF_NBF_df)

#volcano plot
K5_BF_NBF_volc <- ggplot(data = resK5_BF_NBF_df, aes(x = log2FoldChange, y = -log10(padj), fill = diffexpressed)) +
  #geom_vline(xintercept = c(-0.6, 0.6), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size=2, alpha=0.9, shape=21) +
  scale_fill_manual(values=c("dodgerblue2", "grey30", "white"), labels = c("Down in BF", "Not significant", "Up in BF")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  #scale_color_manual(values = c("blue")) +  # to set the colours of our variable
  coord_cartesian(ylim = c(0, 12), xlim = c(-30, 30)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = '', #legend_title,
  x = "log<sub>2</sub>FC", y = "-log<sub>10</sub>p-adj") +
  scale_x_continuous(breaks = seq(-30, 30, 10)) + # to customise the breaks in the x axis
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
K5_BF_NBF_volc
#save as K5_BF_NBF_volc as portrait 4x4

#44_BF_NBF
head(res_44BFNBF)
#Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
res44_BF_NBF_df <- as.data.frame(res44_BF_NBF)
head(res44_BF_NBF_df)
res44_BF_NBF_df <- res44_BF_NBF_df %>%
  rownames_to_column(var="query") %>%
  left_join(gene2query)
head(res44_BF_NBF_df)
res44_BF_NBF_df$diffexpressed <- "NO"
# if log2Foldchange > 0 and padj < 0.05, set as "UP"
res44_BF_NBF_df$diffexpressed[res44_BF_NBF_df$log2FoldChange > 0 & res44_BF_NBF_df$padj < 0.05] <- "UP"
# if log2Foldchange < 0 and padj < 0.05, set as "DOWN"
res44_BF_NBF_df$diffexpressed[res44_BF_NBF_df$log2FoldChange < 0 & res44_BF_NBF_df$padj < 0.05] <- "DOWN"

#Volcano plot
r44_BF_NBF_volc <- ggplot(data = res44_BF_NBF_df, aes(x = log2FoldChange, y = -log10(padj), fill = diffexpressed)) +
  #geom_vline(xintercept = c(-0.6, 0.6), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size=2, alpha=0.9, shape=21) +
  scale_fill_manual(values=c("tomato2", "grey30", "white"), labels = c("Down in BF", "Not significant", "Up in BF")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  #scale_color_manual(values = c("blue")) +  # to set the colours of our variable
  coord_cartesian(ylim = c(0, 5), xlim = c(-20, 20)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = '', #legend_title,
  x = "log<sub>2</sub>FC", y = "-log<sub>10</sub>p-adj") +
  scale_x_continuous(breaks = seq(-20, 20, 10)) + # to customise the breaks in the x axis
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
r44_BF_NBF_volc
#saved as 44_BF_NBF_volc as portrait 4x4



#K5 BM BF
head(res_K5BMBF)
summary(resK5_BM_BF, alpha=0.05)
#LFC > 0 (up)       : 123, 0.3%
#LFC < 0 (down)     : 59, 0.15%
# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
resK5_BM_BF_df <- as.data.frame(resK5_BM_BF)
head(resK5_BM_BF_df)
resK5_BM_BF_df <- resK5_BM_BF_df %>%
  rownames_to_column(var="query") %>%
  left_join(gene2query)
head(resK5_BM_BF_df)
resK5_BM_BF_df$diffexpressed <- "NO"
# if log2Foldchange > 0 and padj < 0.05, set as "UP"
resK5_BM_BF_df$diffexpressed[resK5_BM_BF_df$log2FoldChange > 0 & resK5_BM_BF_df$padj < 0.05] <- "UP"
# if log2Foldchange < 0 and padj < 0.05, set as "DOWN"
resK5_BM_BF_df$diffexpressed[resK5_BM_BF_df$log2FoldChange < 0 & resK5_BM_BF_df$padj < 0.05] <- "DOWN"

#Volcano plot
K5_BM_BF_volc <- ggplot(data = resK5_BM_BF_df, aes(x = log2FoldChange, y = -log10(padj), shape = diffexpressed, fill=diffexpressed)) +
  geom_vline(xintercept = 0, col = "gray80", linetype = 'dotted') +
geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
geom_point(size=2, alpha=0.9) +
scale_shape_manual(values=c(25, 21, 24), labels = c("Down in M", "Not significant", "Up in M"))+
scale_fill_manual(values=c("dodgerblue2", "grey30", "dodgerblue2"), labels = c("Down in M", "Not significant", "Up in M")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
#scale_color_manual(values = c("blue")) +  # to set the colours of our variable
coord_cartesian(ylim = c(0, 5), xlim = c(-25, 25)) + # since some genes can have minuslog10padj of inf, we set these limits
labs(color = '', #legend_title,
x = "log<sub>2</sub>FC", y = "-log<sub>10</sub>p-adj") +
scale_x_continuous(breaks = seq(-25, 25, 10)) + # to customise the breaks in the x axis
theme(panel.background = element_rect(color="black", fill="white"),
axis.text = element_text(colour="black", size=10),
axis.title.x = ggtext::element_markdown(face="bold", size=12),
axis.title.y = ggtext::element_markdown(face="bold",size=12),
panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
legend.position="right",
legend.background=element_rect(color="black"),
legend.text=element_text(face="bold", color="black", size=8),
legend.title= element_blank(),
legend.text.align = 0,
legend.key.size=unit(10,"point"))

K5_BM_BF_volc
#saved as K5_BM_BF_volc as portrait 4x4
```

Venn Diagram comparing degs from K5 and 44 BF NBF
```{r}
head(resK5_BF_NBF)
resK5_BF_NBF <- read.delim("./K5_BF_NBF_DE_26.txt")
head(resK5_BF_NBF)
resK5_BF_NBF$lfc_K5 <- resK5_BF_NBF$log2FoldChange
head(resK5_BF_NBF)
res44_BF_NBF <- read.delim("./44_BF_NBF_DE_26.txt")
res44_BF_NBF
res44_BF_NBF$lfc_44 <- res44_BF_NBF$log2FoldChange
head(res44_BF_NBF)
# look at ALL DEGs
K5_DEGs05=row.names(resK5_BF_NBF[resK5_BF_NBF$padj<0.05 & !is.na(resK5_BF_NBF$padj),])
length(resK5_BF_NBF)
length(K5_DEGs05) #135
r44_DEGs05=row.names(res44_BF_NBF[res44_BF_NBF$padj<0.05 & !is.na(res44_BF_NBF$padj),])
length(r44_DEGs05) #23
candidates_K544=list("K5"=K5_DEGs05, "44"=r44_DEGs05)

prettyvenn=venn.diagram(
  x = candidates_K544,
  category.names=c("K5 (135)", "44 (23)"),
  col=c("black","black"),
  lwd=c(3,3),
  filename=NULL,
  #main = "DEGs Under Starvation",
  main.cex = 2,
  main.fontface = "bold",
  main.fontfamily = "sans",
  #col = "transparent",
  fill = c("dodgerblue2","tomato2"),
  alpha = 0.3,
  cex = 2,
  fontfamily = "sans",
  fontface = "bold",
  cat.default.pos = "text",
  cat.col = c("black", "black"),
  cat.cex = 1.5,
  cat.fontfamily = "sans",
  cat.dist = c(0.25, 0.14),
  cat.pos = c(-14.5, -3)
  )
grid.draw(prettyvenn)
#saved as K544DEG_Venn as 5x5
```

read back in GO outputs for comparing GO terms
```{r}
#K5 BFNBF
go_bp_K5BFNBF <- read.csv("./MWU_BP_K5BFNBF_GO.csv", header=TRUE, sep=" ")
head(go_bp_K5BFNBF)
write.table(go_bp_K5BFNBF, "./MWU_BP_K5BFNBF.txt", quote=F, sep="\t", row.names=FALSE)
bp_K5BFNBF = read.csv("./BP_K5BFNBF_GO.csv", sep="\t")
head(bp_K5BFNBF)
colnames(bp_K5BFNBF)[4]="query"

go_bp_K5BFNBF_UP <- go_bp_K5BFNBF %>%
  subset(go_bp_K5BFNBF$p.adj < 0.1 & go_bp_K5BFNBF$delta.rank > 0)
top20_K5BFNBF_UP <- top_n(go_bp_K5BFNBF_UP, -20, p.adj)
go_bp_K5BFNBF_DOWN <- go_bp_K5BFNBF %>%
  subset(go_bp_K5BFNBF$p.adj < 0.1 & go_bp_K5BFNBF$delta.rank < 0)
top20_K5BFNBF_DOWN <- top_n(go_bp_K5BFNBF_DOWN, -20, p.adj)

#44 BFNBF
go_bp_44BFNBF <- read.csv("./MWU_BP_44BFNBF_GO.csv", header=TRUE, sep=" ")
head(go_bp_44BFNBF)
write.table(go_bp_44BFNBF, "./MWU_BP_44BFNBF.txt", quote=F, sep="\t", row.names=FALSE)
bp_44BFNBF = read.csv("./BP_44BFNBF_GO.csv", sep="\t")
head(bp_44BFNBF)
colnames(bp_44BFNBF)[4]="query"

go_bp_44BFNBF_UP <- go_bp_44BFNBF %>%
  subset(go_bp_44BFNBF$p.adj < 0.1 & go_bp_44BFNBF$delta.rank > 0)
top20_44BFNBF_UP <- top_n(go_bp_44BFNBF_UP, -20, p.adj)
go_bp_44BFNBF_DOWN <- go_bp_44BFNBF %>%
  subset(go_bp_44BFNBF$p.adj < 0.1 & go_bp_44BFNBF$delta.rank < 0)
top20_44BFNBF_DOWN <- top_n(go_bp_44BFNBF_DOWN, -20, p.adj)
```

compare the two sets
```{r}
go_bp_44BFNBF_SIG <- go_bp_44BFNBF %>%
  subset(go_bp_44BFNBF$p.adj < 0.1)
go_bp_K5BFNBF_SIG <- go_bp_K5BFNBF %>%
  subset(go_bp_K5BFNBF$p.adj < 0.1)
# look at ALL sig GOs
K5_DEGOs=go_bp_K5BFNBF_SIG$term
length(K5_DEGOs)
#775
r44_DEGOs=go_bp_44BFNBF_SIG$term
length(r44_DEGOs)
#551
candidates_K544GO=list("K5"=K5_DEGOs, "44"=r44_DEGOs)

prettyvenn=venn.diagram(
  x = candidates_K544GO,
  category.names=c("K5 (775)", "44 (551)"),
  col=c("black","black"),
  lwd=c(3,3),
  filename=NULL,
  #main = "DEGs Under Starvation",
  main.cex = 2,
  main.fontface = "bold",
  main.fontfamily = "sans",
  #col = "transparent",
  fill = c("dodgerblue2","tomato2"),
  alpha = 0.3,
  cex = 2,
  fontfamily = "sans",
  fontface = "bold",
  cat.default.pos = "text",
  cat.col = c("black", "black"),
  cat.cex = 1.5,cat.fontfamily = "sans",
  cat.dist = c(0.22, 0.20),
  cat.pos = c(-14.5, -3)
  )
grid.draw(prettyvenn) #saved as K544GO_Venn
```

look at the shared GOs
```{r}
intersect05 <- intersect(K5_DEGOs, r44_DEGOs)

#install.packages("ggpubr")
library(ggpubr)

head(go_bp_K5BFNBF_SIG)
shared_K5 <- as.data.frame(go_bp_K5BFNBF_SIG)[go_bp_K5BFNBF_SIG$term %in% intersect05, ] %>% select(delta.rank, term)
k5_name <- c("K5_delta_rank", "term")
colnames(shared_K5) <- k5_name
shared_44 <- as.data.frame(go_bp_44BFNBF_SIG)[go_bp_44BFNBF_SIG$term %in% intersect05, ] %>% select(delta.rank, term)
r44_name <- c("r44_delta_rank", "term")
colnames(shared_44) <- r44_name
shared_both_delta <- full_join(shared_K5, shared_44)
View(shared_both_delta)
head(shared_both_delta)

GO_scatter2 <- ggscatter(shared_both_delta, x = "K5_delta_rank", y = "r44_delta_rank",
  add = "reg.line", conf.int = TRUE,
  cor.coef = TRUE, cor.method = "pearson",
  xlab = "GO Delta Rank - K5", ylab = "GO Delta Rank - 44")+ theme(panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"))
#saved as shared_DEGO_correlation as 5x5

head(shared_both_delta)
shared_both_delta_name <- shared_both_delta %>%
  left_join(go_bp_44BFNBF_SIG) %>%
  select(-K5_delta_rank)
```

Make barplots with the top20 up and down
```{r}
head(top20_K5BFNBF_UP)
top20_K5BFNBF_UP_plot <- ggplot(top20_K5BFNBF_UP, aes(x=name, y=delta.rank)) +
  geom_bar(stat = "identity", fill = "white", color = "dodgerblue2", alpha = 0.75 ) +
  #scale_fill_manual(values = "white" ) +
  #scale_color_manual(values = "dodgerblue2") +
  coord_flip( ) +
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
top20_K5BFNBF_UP_plot
#top20_K5BFNBF_UP_plot as 6x4

top20_K5BFNBF_DOWN_plot <- ggplot(top20_K5BFNBF_DOWN, aes(x=name, y=delta.rank)) +
  geom_bar(stat = "identity", fill = "dodgerblue", color = "dodgerblue2", alpha=0.75 ) +
  #scale_fill_manual(values = "white" ) +
  #scale_color_manual(values = "dodgerblue2") +
  coord_flip( ) +
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
top20_K5BFNBF_DOWN_plot
#top20_K5BFNBF_DOWN_plot

top20_44BFNBF_UP_plot <- ggplot(top20_44BFNBF_UP, aes(x=name, y=delta.rank)) +
  geom_bar(stat = "identity", fill = "white", color = "tomato2", alpha = 0.75 ) +
  #scale_fill_manual(values = "white" ) +
  #scale_color_manual(values = "tomato2") +
  coord_flip( ) +
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
top20_44BFNBF_UP_plot
#top20_44BFNBF_UP_plot

top20_44BFNBF_DOWN_plot <- ggplot(top20_44BFNBF_DOWN, aes(x=name, y=delta.rank)) +
  geom_bar(stat = "identity", fill = "tomato2", color = "tomato2", alpha=0.75 ) +
  #scale_fill_manual(values = "white" ) +
  #scale_color_manual(values = "tomato") +
  coord_flip( ) +
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
top20_44BFNBF_DOWN_plot
#top20_44BFNBF_DOWN_plot
```
return to big picture comparisons to add up and down DEGs markers to make volcano plots
```{r}
#all M to F
head(res_MF)
#volcano plot
# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
res_M_F_df <- as.data.frame(res_M_F)
head(res_M_F_df)
res_M_F_df <- res_M_F_df %>%
  rownames_to_column(var="query") %>%
  left_join(gene2query)
head(res_M_F_df)
res_M_F_df$diffexpressed <- "NO"
# if log2Foldchange > 0 and padj < 0.05, set as "UP"
res_M_F_df$diffexpressed[res_M_F_df$log2FoldChange > 0 & res_M_F_df$padj < 0.05] <- "UP"
# if log2Foldchange < 0 and padj < 0.05, set as "DOWN"
res_M_F_df$diffexpressed[res_M_F_df$log2FoldChange < 0 & res_M_F_df$padj < 0.05] <- "DOWN"
head(res_M_F_df)
res_MF_05 <- as.data.frame(res_MF[which(res_MF$padj < 0.05), ])

M_F_volc <- ggplot(data = res_M_F_df, aes(x = log2FoldChange, y = -log10(padj), shape = diffexpressed, fill=diffexpressed)) +
  geom_vline(xintercept = 0, col = "gray80", linetype = 'dotted') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size=2, alpha=0.9) +
  scale_shape_manual(values=c(25, 21, 24), labels = c("Down in M", "Not significant", "Up in M"))+
  scale_fill_manual(values=c("orchid3", "grey30", "orchid3"), labels = c("Down in M", "Not significant", "Up in M")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  #scale_color_manual(values = c("blue")) +  # to set the colours of our variable
  coord_cartesian(ylim = c(0, 6), xlim = c(-25, 25)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = '', #legend_title,
  x = "log<sub>2</sub>FC", y = "-log<sub>10</sub>p-adj") +
  scale_x_continuous(breaks = seq(-25, 25, 10)) + # to customise the breaks in the x axis
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
M_F_volc
#M_F_volc as 6 x 4

#all K5 to 44
head(res_K5_44)
#volcano plot
# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
res_K5_44_df <- as.data.frame(res_K5_44)
head(res_K5_44_df)
res_K5_44_df <- res_K5_44_df %>%
  rownames_to_column(var="query") %>%
  left_join(gene2query)
head(res_K5_44_df)
res_K5_44_df$diffexpressed <- "NO"
# if log2Foldchange > 0 and padj < 0.05, set as "UP"
res_K5_44_df$diffexpressed[res_K5_44_df$log2FoldChange > 0 & res_K5_44_df$padj < 0.05] <- "UP"
# if log2Foldchange < 0 and padj < 0.05, set as "DOWN"
res_K5_44_df$diffexpressed[res_K5_44_df$log2FoldChange < 0 & res_K5_44_df$padj < 0.05] <- "DOWN"
head(res_K5_44_df)
res_K544_05 <- as.data.frame(res_K544[which(res_K544$padj < 0.05), ])

K5_44_volc <- ggplot(data = res_K5_44_df, aes(x = log2FoldChange, y = -log10(padj), shape = diffexpressed, fill=diffexpressed)) +
  geom_vline(xintercept = 0, col = "gray80", linetype = 'dotted') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size=2, alpha=0.9, shape=21) +
  #scale_shape_manual(values=c(25, 21, 24), labels = c("Down in K5", "Not significant", "Up in K5"))+
  scale_fill_manual(values=c("tomato2", "grey30", "dodgerblue2"), labels = c("Down in K5", "Not significant", "Up in K5")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  #scale_color_manual(values = c("blue")) +  # to set the colours of our variable
  coord_cartesian(ylim = c(0, 6), xlim = c(-10, 10)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = '', #legend_title,
  x = "log<sub>2</sub>FC", y = "-log<sub>10</sub>p-adj") +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customise the breaks in the x axis
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
K5_44_volc
#K5_44_volc 6x4

#all Bleach to all Non-Bleach
head(res_BNB)
#volcano plot
# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
res_B_NB_df <- as.data.frame(res_B_NB)
head(res_B_NB_df)
res_B_NB_df <- res_B_NB_df %>%
  rownames_to_column(var="query") %>%
  left_join(gene2query)
head(res_B_NB_df)
res_B_NB_df$diffexpressed <- "NO"
# if log2Foldchange > 0 and padj < 0.05, set as "UP"
res_B_NB_df$diffexpressed[res_B_NB_df$log2FoldChange > 0 & res_B_NB_df$padj < 0.05] <- "UP"
# if log2Foldchange < 0 and padj < 0.05, set as "DOWN"
res_B_NB_df$diffexpressed[res_B_NB_df$log2FoldChange < 0 & res_B_NB_df$padj < 0.05] <- "DOWN"
head(res_B_NB_df)

B_NB_volc <- ggplot(data = res_B_NB_df, aes(x = log2FoldChange, y = -log10(padj), shape = diffexpressed, fill=diffexpressed)) +
  geom_vline(xintercept = 0, col = "gray80", linetype = 'dotted') +
  geom_hline(yintercept = -log10(0.05), col = "gray", linetype = 'dashed') +
  geom_point(size=2, alpha=0.9, shape=21) +
  #scale_shape_manual(values=c(25, 21, 24), labels = c("Downregulated", "Not significant", "Upregulated"))+
  scale_fill_manual(values=c("orchid3", "grey30", "white"), labels = c("Down in B", "Not significant", "Up in B")) + # to set the labels in case we want to overwrite the categories from the dataframe (UP, DOWN, NO)
  #scale_color_manual(values = c("blue")) +  # to set the colours of our variable
  coord_cartesian(ylim = c(0, 4), xlim = c(-10, 10)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = '', #legend_title,
  x = "log<sub>2</sub>FC", y = "-log<sub>10</sub>p-adj") +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customise the breaks in the x axis
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
B_NB_volc
#B_NB_volc 6x4
```

reading in GO outputs for K5BMBF for top20
```{r}
go_bp_K5BMBF <- read.csv("./MWU_BP_K5BMBF_GO.csv", header=TRUE, sep=" ")
write.table(go_bp_K5BMBF, "./MWU_BP_K5BMBF.txt", quote=F, sep="\t", row.names=FALSE)
bp_K5BMBF = read.csv("./BP_K5BMBF_GO.csv", sep="\t")
head(bp_K5BMBF)
colnames(bp_K5BMBF)[4]="query"
go_bp_K5BMBF_SIG <- go_bp_K5BMBF %>%
  subset(go_bp_K5BMBF$p.adj < 0.1)

go_bp_K5BMBF_UP <- go_bp_K5BMBF %>%
  subset(go_bp_K5BMBF$p.adj < 0.1 & go_bp_K5BMBF$delta.rank > 0)
top20_K5BMBF_UP <- top_n(go_bp_K5BMBF_UP, -20, p.adj)

go_bp_K5BMBF_DOWN <- go_bp_K5BMBF %>%
  subset(go_bp_K5BMBF$p.adj < 0.1 & go_bp_K5BMBF$delta.rank < 0)
top20_K5BMBF_DOWN <- top_n(go_bp_K5BMBF_DOWN, -20, p.adj)

#install.packages("ggpattern")
library(ggpattern)

top20_K5BMBF_UP_plot <- ggplot(top20_K5BMBF_UP, aes(x=name, y=delta.rank)) +
  #geom_bar(stat = "identity", fill = "white", color = "dodgerblue2", alpha = 0.75 ) +
  geom_col_pattern(pattern = "pch", pattern_shape = 24, pattern_angle = 0, pattern_color = "dodgerblue2", stat = "identity", fill = "white", color = "dodgerblue2", alpha = 0.75 )+
  #scale_fill_manual(values = "white" ) +
  #scale_color_manual(values = "dodgerblue2") +
  coord_flip( ) +
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
top20_K5BMBF_UP_plot #saved

top20_K5BMBF_DOWN_plot <- ggplot(top20_K5BMBF_DOWN, aes(x=name, y=delta.rank)) +
  #geom_bar(stat = "identity", fill = "white", color = "dodgerblue2", alpha = 0.75 ) +
  geom_col_pattern(pattern = "pch", pattern_shape = 25, pattern_angle = 0, pattern_color = "dodgerblue2", stat = "identity", fill = "white", color = "dodgerblue2", alpha = 0.75 )+
  #scale_fill_manual(values = "white" ) +
  #scale_color_manual(values = "dodgerblue2") +
  coord_flip( ) +
  theme(panel.background = element_rect(color="black", fill="white"),
  axis.text = element_text(colour="black", size=10),
  axis.title.x = ggtext::element_markdown(face="bold", size=12),
  axis.title.y = ggtext::element_markdown(face="bold",size=12),
  panel.grid.major = element_line(colour = "gray80", linewidth = 0.5, linetype="dotted"),
  legend.position="right",
  legend.background=element_rect(color="black"),
  legend.text=element_text(face="bold", color="black", size=8),
  legend.title= element_blank(),
  legend.text.align = 0,
  legend.key.size=unit(10,"point"))
top20_K5BMBF_DOWN_plot #saved
```

pulling DEGs for each portion of GO venn
```{r}
bp_K5BFNBF_q2term2name <- bp_K5BFNBF[,c(1,2,4)] %>%
  subset(term %in% go_bp_K5BFNBF_SIG$term)
length(unique(bp_K5BFNBF_q2term2name$query))
length(unique(bp_K5BFNBF_q2term2name$term))
#775
length(unique(bp_K5BFNBF_q2term2name$query))
#9510

head(shared_both_delta)
bp_K5BFNBF_q2term2nameSHARED <- bp_K5BFNBF[,c(1,2,4)] %>%
  subset(term %in% shared_both_delta$term)
length(unique(bp_K5BFNBF_q2term2name$term))
length(unique(bp_K5BFNBF_q2term2nameSHARED$term))
#212
length(unique(bp_K5BFNBF_q2term2nameSHARED$query))
#7335

bp_44BFNBF_q2term2nameSHARED <- bp_44BFNBF[,c(1,2,4)] %>%
  subset(term %in% shared_both_delta$term)
length(unique(bp_44BFNBF_q2term2nameSHARED$term))
#212
length(unique(bp_44BFNBF_q2term2nameSHARED$query))
#7120
head(bp_44BFNBF_q2term2nameSHARED)

K5_DEGs05_df=resK5_BF_NBF_df[resK5_BF_NBF_df$padj<0.05 & !is.na(resK5_BF_NBF_df$padj),]
K5_DEGs05_df_GO <- K5_DEGs05_df %>%
  subset(query %in% bp_K5BFNBF_q2term2nameSHARED$query)
head(K5_DEGs05_df_GO)
nrow(K5_DEGs05_df_GO)
#20

rld_df <- as.data.frame(rld)
K5_DEGs05_df_shared <- K5_DEGs05_df %>%
  subset(query %in% bp_K5BFNBF_q2term2nameSHARED$query)
head(K5_DEGs05_df_shared)
rld_df_name_K5_shared <- rld_df %>%
  rownames_to_column(var="query") %>%
  right_join(K5_DEGs05_df_shared) %>%
  select(-baseMean, -log2FoldChange, -lfcSE, -stat, -pvalue, -padj, -lfc_K5, -diffexpressed, -query) %>%
  column_to_rownames(var="gene_ID")


##plot the DEGs from K5 that have a shared DEG with 44
#subset out k5
rld_df_name_K5_shared <- rld_df_name_K5_shared[,c(1,6,13,2,9,10)]
k5_shared_means = apply(rld_df_name_K5_shared, 1, mean)
explc_k5_shared = rld_df_name_K5_shared-k5_shared_means

k5_shared_hm <- pheatmap(as.matrix(explc_k5_shared),
  color=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(20)),
  cluster_rows = T, show_rownames = T, fontsize_row=8, fontsize_col=10, treeheight_row = 10, treeheight_col = 10,
  cluster_cols = F, show_colnames = T, angle_col= "315", cellheight = 8, cellwidth = 14,
  breaks=c(seq(from=-2,to=2,by=0.2)),
  scale="none",
  legend=T)

#same for 44
bp_44BFNBF_q2term2nameSHARED <- bp_44BFNBF[,c(1,2,4)] %>%
  subset(term %in% shared_both_delta$term)
length(unique(bp_44BFNBF_q2term2nameSHARED$term))
#212
length(unique(bp_44BFNBF_q2term2nameSHARED$query))
#7120
r44_DEGs05_df=res44_BF_NBF_df[res44_BF_NBF_df$padj<0.05 & !is.na(res44_BF_NBF_df$padj),]
r44_DEGs05_df_shared <- r44_DEGs05_df %>%
  subset(query %in% bp_44BFNBF_q2term2nameSHARED$query)
head(r44_DEGs05_df_shared)
nrow(r44_DEGs05_df_shared)
#5
rld_df_name_44_shared <- rld_df %>%
  rownames_to_column(var="query") %>%
  right_join(r44_DEGs05_df_shared) %>%
  select(-baseMean, -log2FoldChange, -lfcSE, -stat, -pvalue, -padj, -lfc_44, -diffexpressed, -query) %>%
  column_to_rownames(var="gene_ID")
##plot the DEGs from 44 that have a shared GO with K5
#subset out 44
rld_df_name_44_shared <- rld_df_name_44_shared[,c(3,11,12,4,5,15)]
r44_shared_means = apply(rld_df_name_44_shared, 1, mean)
explc_44_shared = rld_df_name_44_shared-r44_shared_means

r44_shared_hm <- pheatmap(as.matrix(explc_44_shared),
  color=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(20)),
  cluster_rows = T, show_rownames = T, fontsize_row=8, fontsize_col=10, treeheight_row = 10, treeheight_col = 10,
  cluster_cols = F, show_colnames = T, angle_col= "315", cellheight = 8, cellwidth = 14,
  breaks=c(seq(from=-2,to=2,by=0.2)),
  scale="none",
  legend=T)
```

now get the uniques
```{r}
#k5
bp_K5BFNBF_SIG <- bp_K5BFNBF %>%
  subset(term %in% go_bp_K5BFNBF_SIG$term)
length(unique(bp_K5BFNBF_SIG$term))
#775
bp_K5BFNBF_q2term2nameUNIQUE <- bp_K5BFNBF_SIG[,c(1,2,4)] %>%
  subset(!(term %in% shared_both_delta$term))
length(unique(bp_K5BFNBF_q2term2nameUNIQUE$term))
#563

K5_DEGs05_df_unique <- K5_DEGs05_df %>%
  subset(query %in% bp_K5BFNBF_q2term2nameUNIQUE$query)
nrow(K5_DEGs05_df_unique)
#26

rld_df_name_K5_unique <- rld_df %>%
  rownames_to_column(var="query") %>%
  right_join(K5_DEGs05_df_unique) %>%
  select(-baseMean, -log2FoldChange, -lfcSE, -stat, -pvalue, -padj, -lfc_K5, -diffexpressed, -query) %>%
  column_to_rownames(var="gene_ID")
head(rld_df_name_K5_unique)

#subset out k5
rld_df_name_K5_unique <- rld_df_name_K5_unique[,c(1,6,13,2,9,10)]
k5_unique_means = apply(rld_df_name_K5_unique, 1, mean)
explc_k5_unique = rld_df_name_K5_unique-k5_unique_means

k5_unique_hm <- pheatmap(as.matrix(explc_k5_unique),
  color=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(20)),
  cluster_rows = T, show_rownames = T, fontsize_row=8, fontsize_col=10, treeheight_row = 10, treeheight_col = 10,
  cluster_cols = F, show_colnames = T, angle_col= "315", cellheight = 8, cellwidth = 14,
  breaks=c(seq(from=-2,to=2,by=0.2)),
  scale="none",
  legend=T)

bp_K5BFNBF_q2term2nameUNIQUE_name <- unique(bp_K5BFNBF_q2term2nameUNIQUE$name)
go_bp_K5BFNBF_SIG_unique <- go_bp_K5BFNBF_SIG %>%
  subset(term %in% bp_K5BFNBF_q2term2nameUNIQUE$term)

#44
bp_44BFNBF_SIG <- bp_44BFNBF %>%
  subset(term %in% go_bp_44BFNBF_SIG$term)
bp_44BFNBF_q2term2nameUNIQUE <- bp_44BFNBF_SIG[,c(1,2,4)] %>%
  subset(!(term %in% shared_both_delta$term))
length(unique(bp_44BFNBF_q2term2nameUNIQUE$term))
#339
length(unique(bp_44BFNBF_q2term2nameUNIQUE$query))

bp_44BFNBF_SIG <- bp_44BFNBF %>%
  subset(term %in% go_bp_44BFNBF_SIG$term)
bp_44BFNBF_q2term2nameUNIQUE <- bp_44BFNBF_SIG[,c(1,2,4)] %>%
  subset(!(term %in% shared_both_delta$term))
length(unique(bp_44BFNBF_q2term2nameUNIQUE$term))
#339
go_bp_44BFNBF_SIG_unique <- go_bp_44BFNBF_SIG %>%
  subset(term %in% bp_44BFNBF_q2term2nameUNIQUE$term)
length(unique(bp_44BFNBF_q2term2nameUNIQUE$query))
nrow(r44_DEGs05_df)
#23
r44_DEGs05_df_unique <- r44_DEGs05_df %>%
  subset(query %in% bp_44BFNBF_q2term2nameUNIQUE$query)
head(r44_DEGs05_df_unique)
nrow(r44_DEGs05_df_unique)
#4
rld_df_name_44_unique <- rld_df %>%
  rownames_to_column(var="query") %>%
  right_join(r44_DEGs05_df_unique) %>%
  select(-baseMean, -log2FoldChange, -lfcSE, -stat, -pvalue, -padj, -lfc_44, -diffexpressed, -query) %>%
  column_to_rownames(var="gene_ID")

##plot!
#subset out 44
rld_df_name_44_unique <- rld_df_name_44_unique[,c(3,11,12,4,5,15)]
r44_unique_means = apply(rld_df_name_44_unique, 1, mean)
explc_44_unique = rld_df_name_44_unique-r44_unique_means
r44_unique_hm <- pheatmap(as.matrix(explc_44_unique),
  color=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(20)),
  cluster_rows = T, show_rownames = T, fontsize_row=8, fontsize_col=10, treeheight_row = 10, treeheight_col = 10,
  cluster_cols = F, show_colnames = T, angle_col= "315", cellheight = 8, cellwidth = 14,
  breaks=c(seq(from=-2,to=2,by=0.2)),
  scale="none",
  legend=T)
```



